---
title: "Microtubule dynamics simulation from experimental data"
output: 
  html_document:
    highlight: "tango"
    df_print: tibble
    toc: yes
---


#### platform       x86_64-apple-darwin15.6.0   
#### arch           x86_64                      
#### os             darwin15.6.0                
#### system         x86_64, darwin15.6.0        
#### status                                     
#### major          3                           
#### minor          6.0                         
#### year           2019                        
#### month          04                          
#### day            26                          
#### svn rev        76424                       
#### language       R                           
#### version.string R version 3.6.0 (2019-04-26)
#### nickname       Planting of a Tree          


```{r, message=FALSE}
library(tidyverse)
```

```{r}

rlnorm2 <- function(n=1, mean, sd) {
  sdlog <- sqrt(log((sd/mean)^2 + 1))
  meanlog <- log(mean) - (sdlog^2) / 2
  rlnorm(n, meanlog = meanlog, sdlog = sdlog)
}
```



### Assume that growth and shrinkage rate follow log-normal distribution.
```{r}
growth_rate <- function(mean, sd){
  rlnorm2(n=1, mean, sd)
  }# µm/sec
shrinkage_rate <- function(mean, sd){
  rlnorm2(n=1, mean, sd)
  }# µm/sec
```

### A function making a single MT dynamics.
```{r}
make_instability <- function(time,
                             catastrophe_freq, 
                             rescue_freq,
                             mean_gr, sd_gr,
                             mean_sr, sd_sr){
  
  l <- list(
    time_s = 1:time,
    state = rep(NA, time),
    velocity = rep(NA, time)
    )
  
  
  ## Make random numbers for growth and shrinkage rate
  gr_init <- growth_rate(mean = mean_gr, sd = sd_gr)
  sr_init <- shrinkage_rate(mean = mean_sr, sd = sd_sr)
  
  # Start with growth
  l$state[[1]] <- "growth"
  l$velocity[[1]] <- gr_init
  
  
  
  for (i in 2:time){
  
    gr <- gr_init
    sr <- sr_init
    
    if (l$state[[i-1]] == "growth") {
      
      l$state[[i]] <- sample(x = c("growth", "shrinkage"), 
                           size = 1, 
                           replace= TRUE, 
                           prob = c(1-catastrophe_freq, catastrophe_freq)
                           )
      # Make a new random number when the phase (shrinkage to growth) is changed.
      sr <- shrinkage_rate(mean = mean_sr, sd = sd_sr) 
      l$velocity[[i]] <- ifelse(l$state[[i]] == "growth", gr, -sr)
      
    }else{
    
      l$state[[i]] <- sample(x = c("growth", "shrinkage"), 
                         size = 1,
                         replace = TRUE,
                         prob = c(rescue_freq, 1-rescue_freq)
                         )
      # Make a new random number when the phase (shrinkage to growth) is changed.
      gr <- growth_rate(mean = mean_gr, sd = sd_gr)
      l$velocity[[i]] <- ifelse(l$state[[i]] == "growth", gr, -sr)
    }
  }
  return(as_tibble(l))
}
```




### A function making MT dynamics with MT length.
```{r}
sim_nMT <- function(group_name = group_name,
                    time,
                    cat_freq,
                    res_freq,
                    mean_gr, sd_gr,
                    mean_sr, sd_sr){
    
  make_instability(
                 time = time,
                 catastrophe_freq = cat_freq,
                 rescue_freq = res_freq,
                 mean_gr = mean_gr, sd_gr = sd_gr,
                 mean_sr = mean_sr, sd_sr = sd_sr
                 ) %>% 
    mutate(
      group = group_name,
      length = accumulate(.$velocity, .f = `+`)
      )
}
```


### Assign parameters gained in the experiments.
```{r}
param <- tibble(
  group_name = c("Control", "Kinesin-13 KO"),
  time = rep(240, length(group_name)), # time duration 240 s, 4 min
  cat_freq = c(0.0093, 0.0022),
  res_freq = c(0.017, 0.036),
  mean_gr = c(0.15, 0.09),
  sd_gr =  c(0.03, 0.01),
  mean_sr = c(0.25, 0.43),
  sd_sr = c(0.06, 0.11)
  )
```

### A function making multiple MT dynamics
```{r}
MT_track <-function(data, nMT){
  map(1:nMT, 
      ~data %>% 
        pmap(sim_nMT) %>% 
        map_dfr(bind_rows) %>% 
        mutate(MT_id = ..1)
      ) %>% 
    map_dfr(bind_rows)
}
```


### Draw MT behavior with 20 examples.
```{r}
set.seed(123)
g <- 
  MT_track(data = param, nMT = 20) %>% 
  ggplot(aes(x = time_s, y = length, color = group)) + 
  geom_path()+
  theme(axis.text=element_text(size=12)) + 
  labs(x = "Time (s)", y = "Length change (µm)", color = "") + 
  ggtitle("Simulated microtubule dynamics")+
  theme(axis.title.x=element_text(lineheight=.9,family="Helvetica",face="bold"),
        axis.title.y=element_text(lineheight=.9,family="Helvetica",face="bold"),
        plot.title=element_text(size=rel(1.4), lineheight=.9,family="Helvetica",face="bold", 
                                hjust = 0.5))


g
```


### Simulation with 4000 MTs
```{r}
set.seed(123)
df <- 
  MT_track(data = param, nMT = 4000) %>%
  group_by(group, MT_id) %>% 
  summarise(mean_length = mean(length),
            final_length = last(length))
```

### Calculation of mean of length change 
```{r}
df %>% 
  group_by(group) %>% 
  summarise(
    mean_mean_length = mean(mean_length),
    mean_final_length = mean(final_length)
            )
```

### Visualization of mean MT length changes in 4 min.
```{r}
df %>% 
  ggplot(aes(x = group,  y = mean_length)) + 
  ggbeeswarm::geom_quasirandom() + 
  theme(axis.text=element_text(size=12)) + 
  labs(x = "", y = "Mean length change (µm)", color = "") + 
  ggtitle("Simulated mean microtubule length change in 4 min")+
  theme(axis.title.x=element_text(lineheight=.9,family="Helvetica",face="bold"),
        axis.title.y=element_text(lineheight=.9,family="Helvetica",face="bold"),
        plot.title=element_text(size=rel(1.4), lineheight=.9,family="Helvetica",face="bold", 
                                hjust = 0.5)) + 
  ylim(c(-22, 22)) # removed 3 outliers in 4000 to make it easy to see
```
### Visualization of MT length change in 4 min elongation.
```{r}
df %>% 
  ggplot(aes(x = group,  y = final_length)) + 
  ggbeeswarm::geom_quasirandom()+
  labs(x = "", y = "Length change (µm)", color = "") + 
  ggtitle("Simulated microtubule length change in 4 min")+
  theme(axis.text=element_text(size=12)) + 
  theme(axis.title.x=element_text(lineheight=.9,family="Helvetica",face="bold"),
        axis.title.y=element_text(lineheight=.9,family="Helvetica",face="bold"),
        plot.title=element_text(size=rel(1.4), lineheight=.9,family="Helvetica",face="bold", 
                                hjust = 0.5)) + 
  ylim(c(-30, 40)) # removed 6 outliers in 4000 to make it easy to see
```


### Mutate simulated data for calculating probability of becoming longer MTs in Control.
```{r}
df_spread <- 
  df %>%
  select(-mean_length) %>% 
  spread(key = group, 
         value = final_length) %>% 
  mutate(bool = Control >`Kinesin-13 KO`,
         difference = Control-`Kinesin-13 KO`)
```

### Distribution of the difference between Control and KO when simulation is performed.
```{r}
df_spread %>%
  gather(key = group, value = len, Control, `Kinesin-13 KO`) %>% 
  ggplot(aes(x = "", y = difference))+
  ggbeeswarm::geom_quasirandom()
```

### Mean of Control and KO difference in microtubule length change 
### and probability of Control microtubules becoming longer than KO
```{r}
df_spread %>%
  summarise(prob = mean(bool), mean_diff = mean(difference))
```



### A function checking whether the number (n) of microtubules in the simulation is sufficient.
```{r}
sampling_test <- function(size, df){
  sample_n(df, size = size, replace = FALSE) %>% 
    summarise_at(vars(bool, difference), mean) %>% 
    rename(prob = bool, mean_diff = difference) %>% 
    mutate(n = size)
}
```

### Execute the function
```{r}
n <- df_spread %>% nrow()
set.seed(123)
sampling_df <- 
  map_df(1:n, 
       ~sampling_test(..1, df = df_spread))
```


### Check if the probability and mean difference converge as n increases
```{r}
# Probability
sampling_df %>% 
  ggplot(aes(x = n, y = prob)) + 
  geom_path()
```

```{r}
# Differnce
sampling_df %>% 
  ggplot(aes(x = n, y = mean_diff)) + 
  geom_path()
```


